---
title: "Chapter 1: Introduction to  R"
output: 
  learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library('leaflet')
library('learnr')
library('checkr')
```

## Arithmetic

### R is an overgrown calculator.

Type the following in the console and hit <kbd>`Enter`</kbd> to *run* these commands one at a time.

``` r
1+1
4-3
3*7
8/3
2^3
pi^2
```

```{r calculatorexcerices1, exercise = TRUE}
    
```

There are many built in functions in R such as `log` or `exp`. Using these functions in R is not much different than using them on your calculator. The function `log` has two arguments, one is required and one is set to a default, $log_{b}(x)$ where *b* is the base. The default is *b* which is set to `exp(1)` (i.e. the natural logarithm).

::: blue
### You try it

Below are some examples of built in functions in R. Run them from your console.

``` r
exp(2)
log(8)
log(8,base=2)
```

```{r youtryit1, exercise = TRUE}
    
```
:::

Now lets try a more complicated equation.

```{r calculatorexample_error1, exercise = TRUE, exercise.eval = FALSE,error=TRUE}
2 + 5*(8^3)- 3*log10)
```

::: {#calculatorexample_error1-hint}
Uh oh, we got an Error. Nothing to worry about, errors happen all the time. Put a open parenthesis `(` before `log10` to fix it and try again.
:::

### Storing results

Most of the time we will want to save our results. To do so we use the *assignment operator* `<-` notation. This lets us save a value into an *object*, which we can then use later similar to a variable $x$ in algebra.

```{r objectrexcerices1, exercise = TRUE}
 height <- 62
```

Assigned objects aren't printed automatically, you must either enter the name of the object, or wrap the command in parenthesis.

```{r objectrexcerices2, exercise = TRUE}
(height <- 62)
height
```

**END**

::: yellow
**Naming Conventions** Be creative, yet informative with your variable names. You will be writing code a lot this semester, and you want your code to be unique to you and not look like a carbon copy of your neighbor. While `pants` and `nifty` are valid variable names, they may not be the best to describe the results from a die roll.
:::

::: blue
### You try it

Run each of the following commands in your console one at a time. Print out the value of `height` after each time. What happens?

``` r
(height <- height + 2)
(height <- 3 * height)
```

```{r youtryit2, exercise = TRUE}

```
:::

## Vectors

A vector is a list of values in order to be able to work with them. For us they will usually represent data collected on a characteristic of the population. In general, we want to give the vector a name so that we can call it later when needed. 

* Create a vector of the first 5 primes, and print this vector.
```{r}
(primes <- c(2,3,5,7,11))
```

------------------------------------------------------------------------
