# Section 2.4: Counting Arguments

This section presents some common methods for counting the number of outcomes in a set.  When there are a lot of outcomes in an experiment, it is convenient to have a method of determining how many outcomes there are in $S$.

\vspace{-0.5cm}

## Multiplication rule: 

Suppose that an experiment has two parts or phases.  In the first part there are $n_{1}$ outcomes and in the second part there are $n_{2}$ outcomes.  The composite experiment which consists of  both parts of the experiment then has $n_{1}\times n_{2}$ possible outcomes.

\vspace{-0.5cm}

### Example: Despite all my rage...
Let $E_{1}$ denote the selection of a rat form a cage containing one female (F) rat and one male (M) rat.  Let $E_{2}$ denote the administering of either drug A, drug B, or a placebo to the selected rat.

* How many possible outcomes are there? 
* List the possible outcomes: 

\vspace{2cm}

Another way of illustrating the multiplication principle is with a *tree diagram*.  The diagram shows that there are $n_{1}$=2 possibilities for the gender of the rat and that for each of these outcomes there are $n_{2}=3$ possibilities for the drug.

\vspace{3cm}

The multiplication rule can be extended to more than two experiments.

\newpage

### You try it
1. Each year starts on one of the seven days (Sunday through Saturday).  Each year is either a leap year (i.e., it includes February 29) or not.  How many different calendars are possible for a year?

\vspace{2cm}


2. A chemical engineer wishes to observe the effects of temperature, pressure, and catalyst concentration on the yield resulting from a certain reaction.  If she intends to include two different temperatures, three pressures, and two levels of catalyst, how many different runs must she make in order to observe each temperature-pressure catalyst combination exactly twice?

\vspace{2cm}


3. A restaurant offers a choice of four appetizers, fourteen entrees, six desserts, and five beverages. How many different meals are possible if a diner intends to order only three items, one from each menu? That is, you can't have two desserts and no entree.



\newpage

## Permutations

* An ordered arrangements of a countable set of objects is called a **permutation**. 
* The number of permutations of $n$ distinct objects is $n!$. 
* The `!` is a function called a **factorial** and is defined as 

$$
n! = n * (n-1) * (n-2) * \ldots * 1
$$

```{r}
factorial(3)
```

### Example

The "ice cream club" is hosting a make-your-own sundae at which the following are provided:

* Ice Cream flavors: Chocolate, Cookies-n-cream, Strawberry, Vanilla
* Toppings: Caramel, Hot Fudge Marshmallow, M&Ms, Nuts, Strawberries

How many different sundaes are possible using one flavor of ice cream and three different toppings?

\vspace{2cm}

How many sundaes are possible using one flavor of ice cream and from 0 to 6 toppings?

\vspace{2cm}


### You try it

There are 9 presidential candidates at a debate. How many different ways can candidates be lined up?


\newpage

## Combinations

If the order of objects is not important, then the number of ways of choosing $k$ distinct objects from a set of $n$ is given by

$$
\binom{n}{k}=\frac{n!}{k!(n-k)!}.
$$

```{r}
choose(3,2)
```



### Example: $\alpha\beta\zeta$

The Alpha Beta Zeta sorority is trying to fill a pledge class of nine new members during fall rush.  Among the twenty-five available candidates, fifteen have been judged marginally acceptable and ten highly desirable.  How many ways can the pledge class be chosen to give a two-to-one ratio of highly desirable to marginally acceptable candidates?


\newpage

### You try it
For each of these, write the R code used to calculate the answer and the answer itself. 

1. Among the 9 presidential candidates at a debate, 3 are republicans and 6 are democrats. How many different line ups are possible if the only ordering that matters is political party (not name)? 

\vspace{3cm}

2. Nine students, five statistics majors and 4 computer science majors, interview for four summer internships sponsored by Google.
    a. In how many ways can Google choose a set of four interns?\vspace{3cm}
    b. In how many can Google choose 2 stat majors and 2 computer science majors?\vspace{3cm}
    c. How many sets of four can be picked such that not everyone in the set is the same major?\vspace{3cm}


\newpage

## Combinatorial Probability

In the previous section our concern focused on counting the number of ways a given operation, or sequence of operations could be performed.  In this section we want to calculate the probability that a certain combination will occur.

### Example: Gender equality in promotions
Ten equally qualified marketing assistants are candidates for promotion to associate buyer; seven are men and three are women.  If the company intends to promote four of the ten at random, what is the probability that exactly two of the four are women?

```{r, eval=FALSE}
total     <-  
two_women <-  
(prob_two_women <-  )
```

\vspace{2cm}

### Example: Urns and chips
An urn contains twenty chips, numbered 1 through 20.  Two are drawn simultaneously.  What is the probability that the numbers on the two chips will differ by more than 2? _Hint: Calculate the complement and subtract from one._

```{r, eval=FALSE}
total <- ___________
1 - ___________
```

\newpage

### You try it:
1. An apartment building has eight floors.  If seven people get on the elevator on the first floor, what is the probability that they all want to get off on different floors? On the same floor? 

\vspace{8cm}


2. If four dice are rolled, what is the probability that each of the four numbers that appear will be different?

\newpage
Additional notes.
