# Section 2.1: Probability Basics

A goal of Statistics is to describe the real world based on limited observations. Observations are influenced by random, and non-random conditions (e.g. the weather, what you ate for breakfast). Probability is a way to mathematically describe random events. 

### Vocabulary

* Experiment: 

\vspace{1cm}

* Outcome:

\vspace{1cm}

* Sample space: 

\vspace{1cm}

* Event:

\vspace{1cm}

* Trial:

### Examples

* Roll a die and observe the number of dots on the face


* Stop a random person on the street and ask what month they are born


* Suppose a traffic light stays red for 90 seconds each cycle. When driving you arrive at the light and observe the amount of time that you are stopped until the light turns green. 


### You try it:
For each of the following problems, identify the sample space and the events described. 

1. Observe eye color of a group of students.
    - Sample space: 
    - Event student does not have blue eyes: 
2. Number of credits a student can take:
    - Sample space: 
    - Event student takes less than 9 credits:
3. Toss a coin and roll a die.
    - Sample space: 
    - Event that you get tails:
4. A soccer team is in the playoffs. The team will play three games and will either win (`w`) or lose (`l`) each game (assume ties are not allowed).
    - Sample Space:
    - Event that at least 2 games are won: 

## Set Definitions

Let $A$ and $B$ be events in a sample space $S$. Complete the following definitions and write an example of each using context situations above. 

* $A\cap B$ 
* $A\cup B$ 
* The *complement* of $A$ is
* The symbol $\emptyset$ is 
* $A$ and $B$ are **disjoint** or **mutually exclusive** if and only if
* $A\cap B^{c}$ 

_Note that _element_ and _outcome_ can be used interchangeably._



### Venn Diagram
The most common kind of picture to make to describe sample spaces and events within sample spaces is a *Venn Diagram*. A Venn diagram uses overlapping circles or other shapes to illustrate the logical relationships between two or more sets of items.  

\vspace{2cm}

### Example

Say 3 roommates are deciding on a pet. They use a Venn Diagram to determine which pet might be the best pick for them.

* Sidney prefers: cat, bird, hamster,spider, goat.
* Ralph prefers: dog, cat, fish,goat.
* Gilbert prefers: horse, cat, dog, turtle snake,goat,fish

Create a Venn diagram that represents this example.

\vspace{6cm}

What pet should they choose? 

\newpage

### You try it:
A single card is drawn from a standard deck of cards. (Not sure what that looks like? See here: https://en.wikipedia.org/wiki/Standard_52-card_deck) 

\vspace{1cm}

Let $A$ be the event that an ace is selected, and let $B$ be the event that a heart is drawn.

1. Define $A$ and $B$ using set notation. 

\vspace{2cm}

2. Write the event space, and what the following mean in context of a deck of cards. 

* $A\cup B$  \vspace{1cm}
* $A\cap B$ \vspace{1cm}
* $B^{c}$  \vspace{1cm}

## Set operations in `R`

We can also rely on `R` to perform union and intersection calculations. The following functions are used to compute intersections and unions. Each function can only take into consideration 2 vectors. 

* `union`:  Either event, or both. ($\cup$)
* `intersect`: Where do the two events overlap. ($\cap$) 
* `setdiff`: Where the two events do not overlap.

### Example

Lets revisit the deck of cards problem from above: _A single card is drawn from a standard deck of cards. Let $A$ be the event that an ace is selected, and let $B$ be the event that a heart is drawn._

First create the sample space. I am also printing the vector to ensure that what's being created is what is intended. Trust, but verify your code. 

\singlespace
```{r}
(numbers <- rep(c(1:10, "J", "Q", "K", "A"), 4))
(suits <- rep(c("H", "C", "D", "S"), each = 13))
(deck <- paste0(numbers, suits))
```

\vspace{1cm}

Define event A
```{r}
aces <- c("AH", "AC", "AD", "AS") 
```

Define event B
```{r}
(hearts <- paste0(c(1:10, "J", "Q", "K", "A"), "H"))
```

Create a Venn Diagram for events A and B


\newpage

Use R functions to find the following statements. Draw a venn diagram for each and shade the region of interest. 

* $A\cup B$ 
```{r}
(aces.and.hearts <- union(aces, hearts))
```

* $A\cap B$ 
```{r}
(ace.of.hearts <- intersect(aces, hearts))
```

* $B^{c}$ 
```{r}
(no.hearts <- setdiff(deck, hearts))
```

\doublespace

### You try it:

Suppose that one card is to be selected from a deck of 20 cards that contains 10 red cards numbered from 1 to 10 and 10 blue cards numbered from 1 to 10. Let $A$ be the event that a card with an even number is selected, let $B$ be the event that a blue card is selected, and let $C$ be the event that a card with a number less than 5 is selected.

Describe the sample space, draw a Venn diagram, and compute each of the following:

*  $A\cap B\cap C$

\vspace{1cm}

*  $B\cup C^{c}$

\vspace{1cm}

*  $A\cap (B\cup C)$.

\vspace{1cm}

*  $A^{c}\cap B^{c}\cap C^{c}$

\vspace{1cm}



## Definition of Probability

The probability of an event describes the proportion of time we expect the event to occur if we observed the event an infinite number of times. 

Let $S$ be a sample space. A valid *probability* of events $A$ is a number $P(A)$ between 0 and 1 (inclusive), so $0\leq P(A)\leq 1$, that satisfies the following *probability axioms*:

* The probability of 
* Probabilities are countably additive. If $A_{1}$, $A_{2}$,...,$A_{n}$ are disjoint then

\vspace{2cm}


## Probability rules
These are some important rules to memorize that come about as a result of the above axioms. Here are a few, there are more in the textbook. 

Let $A$ and $B$ be events in the sample space $S$.

* The probability of $\null$ is 0. 
* If $A$ and $B$ are disjoint then $P(A\cap B)=P(A)+P(B)$.
* $P(A)=1-P(A^{c})$.
* $P(A\cup B)=P(A)+P(B)-P(A\cap B)$



### Example

1. These rules allow you to manipulate equations to find unknown quantities based on known ones using algebra. Let's use these to show that $P\left(A\bigcap B\right) \geq 1-P(A^{C})-P(B^{C})$ for any two events $A$ and $B$ defined on a sample space $S$.

\vspace{4cm}


2. Events $A$ and $B$ are defined on a sample space $S$ such that $P((A\cup B)^{c})=0.5$ and $P(A\cap B)=0.2$. What is the probability that either $A$ or $B$ but not both will occur?

\vspace{4cm}

### You try it:
If 50 percent of the families in a certain city subscribe to the morning newspaper, 65 percent of the families subscribe to the afternoon newspaper, and 85 percent of the families subscribe to at least one of the two newspapers. Draw a Venn Diagram to represent this situation. 

* What percentage of the families subscribe to both newspapers?

\vspace{1cm}

* What percentage of the families subscribe to only the afternoon paper? 

\vspace{1cm}

* What percentage of the families don't subscribe to any paper? 

\vspace{1cm}

### Example

David Diez was interested in exploring the factors that contribute to an email being flagged as spam by Gmail's system. So they downloaded all their emails for a few months in 2012 and noted certain characteristics such as if it was flagged as spam (0 means no, and 1 means yes), and what size of a number it contained (none, small, or big). A _two-way table_ of emails with these two characteristics are shown below. 

```{r, echo=FALSE}
em <- openintro::email
addmargins(table(em$spam, em$number, dnn=c("Spam", "Size of number")))
```

If you were to randomly select an email from this pool, calculate the following probabilities: 

* It is flagged as spam: 
```{r}
367/3921
```

* It has a big number: 
```{r}
545/3921
```

* It is not flagged as spam and has a small number: 
```{r}
2559/3921
```




